### 2024-11-10 README.Rmd
optimizations + revised documentation

embeddingplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com). Includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g. [GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, and reducing dimensionality.

@@

Functions for similarity and distance metrics are as simple as possible:

```{r}
vec1 <- c(1, 5, 2)
vec2 <- c(4, 2, 2)
vec3 <- c(1, -2, -13)

dot_prod(vec1, vec2)            # dot product
cos_sim(vec1, vec2)             # cosine similarity
euc_dist(vec1, vec2)            # Euclidean distance
anchored_sim(vec1, vec2, vec3)  # projection to an anchored vector
```

### Tidy Workflow

### Quanteda Workflow

>>>>

embeddingplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com). It includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g. [GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, reducing dimensionality, and more.

@@

Functions for similarity and distance metrics are as simple as possible; each one takes in vectors and outputs a scalar.

```{r}
vec1 <- c(1, 5, 2)
vec2 <- c(4, 2, 2)
vec3 <- c(-1, -2, -13)

dot_prod(vec1, vec2)            						# dot product
cos_sim(vec1, vec2)             						# cosine similarity
euc_dist(vec1, vec2)            						# Euclidean distance
anchored_sim(vec1, pos = vec2, neg = vec3)  # projection to an anchored vector
```

### Example Tidy Workflow

Given a tidy dataframe of texts, `embed_docs()` will generate embeddings by averaging the embeddings of words in each text (for more information on why this works well, see [Data Science for Psychology, Chapter 18](https://ds4psych.com/decontextualized-embeddings#sec-embedding-magnitude)).

```{r}
library(dplyr)
valence_df <- tribble(
	~id,        ~text,
	"positive", "happy awesome cool nice",
	"neutral",  "ok fine sure whatever",
	"negative", "sad bad horrible angry"
	)

valence_df <- valence_df |> 
	embed_docs("text", glove_twitter_25d, id_col = "id", .keep_all = TRUE)
valence_df
```

To quantify how good and how intense these texts are, we can compare them to the embeddings for "good" and "intense" using `get_similarities()`. Note that this step requires only a dataframe or tibble with numeric columns; the embeddings can come from any source.

```{r}
good_vec <- predict(glove_twitter_25d, "good")
intense_vec <- predict(glove_twitter_25d, "intense")
valence_quantified <- valence_df |> 
	get_similarities(
		dim_1:dim_25, 
		list(
			good = good_vec, 
			intense = intense_vec
			)
		)
valence_quantified
```
### Example Quanteda Workflow

### Other Functions

#### Reduce Dimensionality

It is sometimes useful to reduce the dimensionality of embeddings. This is done with `reduce_dimensionality()`, which by default performs PCA without column normalization.

```{r}
valence_df_2d <- valence_df |> 
	reduce_dimensionality(dim_1:dim_25, 2)
valence_df_2d
```
#### Normalize or Center

#### Magnitude
